---
title: "Four Parameters Categorical"
format: html
author: "Inam Khan"
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
library(broom)  # for tidy()
library(marginaleffects)
```

```{r}
#| label: prepare-data
#| cache: true

nes_92 <- nes |> 
  filter(year == 1992) |> 
  select(sex, pres_vote) |> 
  drop_na() |> 
  mutate(pres_vote = as.factor(case_when(
    pres_vote == "Democrat" ~ "Clinton",
    pres_vote == "Republican" ~ "Bush",
    pres_vote == "Third Party" ~ "Perot"
  )))
```
```{r}
#| label: fit-model
#| cache: true

fit_nes <- multinom_reg(engine = "nnet") |>
  fit(pres_vote ~ sex, data = nes_92)
```


```{r}
#| label: vote-prob-plot
#| fig-width: 8
#| fig-height: 5

library(ggplot2)

preds <- plot_predictions(fit_nes, by = "sex", type = "prob", draw = FALSE)

ggplot(preds, aes(x = sex, y = estimate, color = sex)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  facet_wrap(~ group, nrow = 1) +
  labs(
  title = "Voting Preferences by Candidate and Sex",
  subtitle = "Clinton had stronger support among women; Perot attracted more men",
  x = NULL,
  y = "Estimated Proportion of Vote",
  color = "Sex",
  caption = "Source: 1992 National Election Studies (NES)"
) +
  scale_color_manual(values = c("Female" = "#E69F00", "Male" = "#56B4E9")) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(margin = margin(b = 10)),
    plot.caption = element_text(size = 10),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
  )
```

Understanding how voter preference varies by gender can help candidates tailor their campaign strategies more effectively. Using data from the 1992 National Election Studies survey, we explore the relationship between voter choice and sex in the U.S. presidential election. We model voter preference as a multinomial function of sex, using a logistic link to estimate the probability of choosing each candidate. A potential limitation of our model is that survey respondents may not fully reflect the broader voting population, which could affect the generalizability of our findings. Our results suggest that the estimated probability of a woman voting for Clinton was about 52.5%, with a 95% confidence interval ranging from 49.2% to 55.8%.


$$
P(Y = k) = \frac{e^{\beta_{k0} + \beta_{k1} X_1 + \beta_{k2} X_2 + \cdots + \beta_{kn} X_n}}{\sum_{j=1}^{K} e^{\beta_{j0} + \beta_{j1} X_1 + \beta_{j2} X_2 + \cdots + \beta_{jn} X_n}}
$$

with $Y \sim \text{Multinomial}(\boldsymbol{\rho})$ where $\boldsymbol{\rho} = (\rho_1, \rho_2, \ldots, \rho_K)$ are the probabilities above.

$$
\hat{P}(\text{pres\_vote} = \text{Clinton}) = \frac{e^{0.455 - 0.255 \cdot \text{sex}_{\text{Male}}}}{1 + e^{0.455 - 0.255 \cdot \text{sex}_{\text{Male}}} + e^{-0.852 + 0.420 \cdot \text{sex}_{\text{Male}}}}
$$

$$
\hat{P}(\text{pres\_vote} = \text{Perot}) = \frac{e^{-0.852 + 0.420 \cdot \text{sex}_{\text{Male}}}}{1 + e^{0.455 - 0.255 \cdot \text{sex}_{\text{Male}}} + e^{-0.852 + 0.420 \cdot \text{sex}_{\text{Male}}}}
$$

$$
\hat{P}(\text{pres\_vote} = \text{Bush}) = 1 - \hat{P}(\text{Clinton}) - \hat{P}(\text{Perot})
$$


```{r}
#| label: tidy-table

tidy(fit_nes, conf.int = TRUE) |>
  select(y.level, term, estimate, conf.low, conf.high) |>
  mutate(across(c(estimate, conf.low, conf.high), round, digits = 3)) |>
  gt() |>
  tab_header(
    title = "Multinomial Logistic Regression: Voter Preference by Sex",
    subtitle = "1992 U.S. Presidential Election (NES Data)"
  ) |>
  cols_label(
    y.level = "Vote Choice (vs. Bush)",
    term = "Term",
    estimate = "Estimate",
    conf.low = "95% CI (Lower)",
    conf.high = "95% CI (Upper)"
  ) |>
  tab_source_note(
    source_note = "Source: National Election Studies (NES) 1992"
  )
```
